---
title: "Stock Market Prediction Using Textual and Numerical Analysis"
author: "Atmik"
date: "3/15/2022"
output:
  html_document: default
---

## Using ARIMA Model to Predict the Movement of Stock Market in the Near Future

Importing Libraries

```{r}
library(tidyverse)
library(tseries)
library(forecast)
library(ggplot2)
library(caTools)
library(rpart)
library(syuzhet)
library(randomForest)
```

Reading the file and Data Preprocessing

```{r}
df <- read.csv(file.choose())
start <- which(rownames(df) == "January-2012")
end <- which(rownames(df) == "December-2019")               
df <- df[start:end,]
df$MonthsNew <- rownames(df)
df <- df[,c(4,6)]
colnames(df) <- c("Close", "Month")
df<- df[,c(2,1)]
rownames(df) <- NULL
df %>% separate(Month, c("Month","Year"), sep = "-" )
```

Converting into Time Series

```{r}
ts <- ts(df$Close, freq = 12, start = c(2012, 1), end = c(2019, 12))
ts
plot(ts)
```

Making the Time Series Stationary and Plotting Different Components of Time Series

```{r}
ts_log <- log(ts)
plot(ts_log)

decomp <- decompose(ts_log)
plot(decomp)
```

Findng the Best Fittting ARIMA Model

```{r}
model <- auto.arima(ts_log)
model
```

Plotting ACF and PACF Graphs for the Best Fittting ARIMA Model

```{r}
acf(model$residuals)
pacf(model$residuals)

Box.test(model$residuals, type = 'Ljung-Box')
adf.test(diff(ts_log))

```

Forecasting and Plotting the Forecast

```{r}
f <- forecast(model, 12)
f
autoplot(f) 
```

## Using Natural Language Processing to Gauge the Effect of News Headlines on Stock Market

Importing the Dataset

```{r}
news <- read.csv(file.choose(),stringsAsFactors = F)
news2 <- news
```

Data Preprocessing

```{r}
news$publish_date <- as.Date(as.character(news$publish_date), "%Y%m%d")
news$headline_category <- NULL
news <- news[news$publish_date > as.Date("2012-01-01") & news$publish_date < as.Date("2019-12-31"),]
news <- separate(news, publish_date, c("Year","Month","Day"), "-")
news$Day <- NULL
news$time <- paste(news$Month, "-", news$Year) 
news$Month <- NULL
news$Year <- NULL
news2 <- news
```

Collapsing the Dataframe into Monthly News

```{r}
news3 <- news %>% group_by(time) %>% summarise(headline_text=paste(headline_text, collapse = " "))
```

Sentiment Analysis of the News Headlines

```{r}
text <- as.character(news3$headline_text)
class(news3$headline_text)
s <- get_nrc_sentiment(text)
head(s)
```

Merging the Stock Closing Index and News Headlines Dataframe

```{r}
final <- data.frame(df$Month, s, df$Close)
final$PrevClose[2:96]<- final$df.Close[1:95]
final$PrevClose[1] <- 15454.92
final$Change <- (final$df.Close - final$PrevClose)*100/final$PrevClose 
final$Change <-   round(final$Change, 2)
final$Class <- ifelse(final$Change > 0, 1 , 0)
head(final,10)
```

Feature Selection, Scaling and Factorization

```{r}
final2 <- final[,c(-1,-12:-14)]
final2[,1:10] <- scale(final2[,1:10])
final2$Class <- factor(final2$Class)
class(final2$Class)
head(final2)
```

Splitting Data into Training and Test Data Sets

```{r}
split <- sample.split(final2$Class, SplitRatio = 0.75)
training <- subset(final2, split == T)
test <- subset(final2, split == F)
colnames(final2)
```

Logistic Regression Classifier

```{r}
lr_classifier <- glm(formula = Class ~ . , family = binomial, data = training )
summary(lr_classifier)

prob <- predict(lr_classifier, type = 'response', newdata = test[-11])
prob
y <- ifelse(prob > 0.5, 1, 0)
y
```

Confusion Matrix for Logistic Regression Classifier

```{r}
lr_cm <- table(test$Class, y)
lr_cm
```

Decision Tree Classifier

```{r}
dtclassifier <- rpart(formula = Class ~ . , data = training)

dt_y <- predict(dtclassifier, newdata = test[-11], type = 'class')
dt_y
```

Confusion Matrix for Decision Tree Classifier

```{r}
cm <- table(test$Class, dt_y)
cm
```

Random Forest Classifier

```{r}
rf_classifier <- randomForest(x=training[-11], y=training$Class, ntree = 300)

rf_y <- predict(rf_classifier, newdata = test[-11])
rf_y

```

Confusion Matrix for Random Forest Classifier

```{r}
cm <- table(test$Class, rf_y)
cm
```

*We see the Random Forest Classifier at 300 trees gives the best output.*

### ------------------------------------------------------------------------------------------------------------------------------------------------------------
